---
title: "Lab 5"
author: "Nathan Snell"
date: "2/23/2021"
output: github_document
---

## Introduction 

In this lab we expand on our knowledge of sorting data by filtering, selecting, and arranging data into tables. With these tables we can then compare data sets using different graphs. New skills such as creating and renaming variables, summarizing datasets, and visualizing variable relationships.

## Load Packages

```{r}
library(tidyverse)
```

## Load Data
```{r}
iris
iris <- as_tibble(iris)
iris
```


##Questions and Answers

>1. Rename each variable so that it is all lower-case and uses an underscore _ instead of a period . in the name (the recommended coding style in the tidyverse style guide). Print the resulting table.
```{r}

iris_new <-
  rename( iris, 
          sepal_length = Sepal.Length, 
          sepal_width = Sepal.Width, 
          petal_length = Petal.Length,
          petal_width = Petal.Width,
          species = Species)
iris_new

```

>2. Convert the four numerical variables from mm to cm by multiplying by 10. Print the resulting table.
```{r}
iris_mm <-
mutate(iris_new,
       sepal_length =sepal_length *10,
       sepal_width =sepal_width *10,
       petal_length =petal_length *10,
       petal_width =petal_width *10)
iris_mm  
```

>3. Calculate sepal area and petal area (area is equal to length multiplied by width). Print a table with only the variables sepal area, petal area, and species.
```{r}
iris_area <- select(iris_mm, petal_length, petal_width)
iris_area

iris_area2 <-
mutate(iris_area, area = petal_length * petal_width)

iris_area2
```

>4. Calculate the following statistics for the entire dataset from the sepal length variable and print the resulting table:
sample size
maximum value
minimum value
range
median
first quartile (q1)
third quartile (q2)
inter-quartile range (iqr)
Calculate the following statistics for each species from the petal width variable and print the resulting table:
```{r}
summary_sepallength <-
   summarise(iris_mm,
             sample_size =n(),
             max =max(sepal_length),
             min =min(sepal_length),
             range = max-min,
             median = median(sepal_length),
             first_quartile =quantile(sepal_length, .25),
             third_quartile =quantile(sepal_length, .75),
             IQR = IQR(sepal_length))
summary_sepallength
```

>5. Calculate the following statistics for each species from the petal width variable and print the resulting table:
sample size
mean
standard deviation
variance
standard error of the mean
approximate 95% confidence interval
```{r}
(species_grouped <- group_by(iris_mm, species))
species_grouped

summary_petalwidth <-
  summarise(species_grouped,
            sample_size =n(),
            mean =mean(petal_width),
            sd =sd(petal_width),
            var =var(petal_width),
            standard_error =mean/sqrt(sample_size),
            conf_upper =mean + 2 * standard_error,
            conf_lower =mean - 2 * standard_error)
summary_petalwidth
```

>6 and 7. Visualize the relationship between petal length and species using a strip plot.Then add the mean and 95% confidence interval for each species

```{r}
ggplot(data =iris_mm)+
  geom_jitter(mapping = aes(y =petal_width, x =species))+
  geom_crossbar(data =summary_petalwidth,
  mapping = aes(x =species, y =mean, ymax =conf_upper, ymin =conf_lower),
  color = "blue")
```


>8. Visualize the relationship between petal length, petal width, and species using a scatterplot. Map the two numerical variables to the x and y axes and map species to the color and shape aesthetics.
```{r}
summary_final <- select(iris_mm, petal_length, petal_width, species)                
ggplot(data=summary_final)+
  geom_point(mapping = aes(x =petal_length, y =petal_width, color =species))


```

## Session Information
```{r session-info}
sessioninfo::session_info()
```
```


